Missing links and the topological robustness of food webs

### 2023-02-10 manuscript/ms_C3_v2.Rmd
updated ms

  1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occuring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that model predicted food webs usually have greater connectance than observed ones. Thus we expect that predicted food webs are more robust than observed ones. This expectation has never, to our knowledge, been tested, nor has the effect size been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links, though which does not predict all links correctly. We did this for 12 different food webs from a wide variety of ecosystems. We used three extinction scenarios: random, most connected, and least connected.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, and this can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found the food webs to be robust by 0.52 units and 0.04 units for the most connected and the random species extinction scenarios respectively while no effect in the least connected species extinction scenario.
  5)  These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, and so it would be interesting to compare the dynamical stability of observed and predicted food webs, as well as the topological stability that we have focused on.

keywords:
  - connectance
  - ABC
  - ADBM
  - food web
  - extinction
  - uncertainty
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage {amsmath}
  - \setlength\parindent{24pt}
  - \usepackage{setspace}\doublespacing
# indent: true
mainfont: Times New Roman
bibliography: bibliography.bib
biblio-style: biblatex
output:
  rticles::arxiv_article:
    keep_tex: true
---

@@

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity, and can lead to food web collapse [@ullah2018]. This food web collapse is due to the cascades of secondary extinctions in a food web because of the primary loss of species, for example due to habitat destruction and climate change [@pimmHumanImpactsRates2006; @thomasComparativeLossesBritish2004; @thomasExtinctionRiskClimate2004]. An example of a secondary extinction is when a consumer goes extinction because its sole resource species goes extinct. Therefore, research focused on cascading secondary extinctions also known as 'community viability analysis' have been performed extensively in the past few decades to quantify how robust are food webs to species extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005]. This research revealed that the rate of collapse of a food web is dependent on its structure and complexity [@dunne2002network; @dunneCascadingExtinctionsCommunity2009].

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems, where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. Primary species loss is considered to be extinction of a species due to causes external to the food web. In research with models, the specific cause or causes are usually not specified. Overharvesting of a population which causes extinction could be considered a driver of a primary extinction. A secondary extinction is an extinction caused by a previous extinction (either primary or secondary). Numerous studies have used a topological criteria for assigning a secondary extinction to a species, i.e., if an extinction is of the last resource species of a consumer species, then the consumer species suffers secondary extinction because it has no resources.

When there are few primary extinctions, and these cause many secondary extinctions, a food web is said to be fragile or not robust. When there are few secondary extinctions, a food web is said to be robust. Several studies have shown that the robustness of the food webs increases with food web connectance. Also, the removal (primary extinction) of the most connected species cause considerably more secondary extinctions than the random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. Simulation studies like these, which investigate the impact of primary extinctions in a food web to quantify robustness based on its topological structure, provide an alternate solution to canonical experiments in natural ecosystems which are not possible or very difficult to conduct [@dunneCascadingExtinctionsCommunity2009].

Along with quantifying food web robustness based on its topological structure, studies such as @williamsEffectsNetworkDynamical2008, @broseAllometricScalingEnhances2006 and @martinezDiversityComplexityPersistence have quantified robustness based on the abundance dynamics of a food web. The topological approach of quantifying a food web robustness only requires the food web structure whereas the dynamical approach not only requires the food web structure but also the temporal dynamics of abundance of species in that food web. For example: @williamsEffectsNetworkDynamical2008 combined models of network structure with models of bioenergetic dynamics to study the role of food web topology and nonlinear dynamics on species coexistence in complex ecological networks.

@@

A key assumption of the observed food webs is that they are very well sampled i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and then do not represent all of the feeding links that can occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort as compared to others, whereas some trophic links remain unobserved because of linkage constraints irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. Previous studies such as @caronAddressingEltonianShortfall and @guptaSimultaneouslyEstimatingFood2022 have shown that the predicted food webs from these models usually have greater connectance than the observed ones. Therefore, one solution to compensate for undersampling is to use a food web model such as the Allometric Diet Breadth Model (ADBM) to predict the missing links, and to then measure the robustness of the predicted food web. The ADBM is a mechanistic model constructed using foraging rules based on the body sizes of prey and predator where trophic interactions satisfying those rules would be predicted by the model which are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links, and also may predict some links that could not, in fact occur.

@@

In our study, we investigate the topological robustness of the ADBM predicted food webs and compare it to that of the observed food webs. We expect the ADBM predicted food webs to be more robust as compared to the observed food webs, and for the greater robustness to be related to the amount by which the ADBM overestimates connectance. We do this by simulating primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of primary species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

<!-- [OP] ## Provide overview of the methods -->

In the upcoming sections, we present a detailed account of the implementation of simulation of primary extinctions for three different extinction scenarios on 12 food webs predicted by the ADBM from a wide variety of ecosystems and compute the resultant secondary extinctions. We then describe a robustness metric of those predicted food webs and how we compare to properties of the food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. We chose this model because we have already used it to predict missing links in the predicted food webs, whereby it consistently overestimated connectance  @guptaSimultaneouslyEstimatingFood2022. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables used in the model are the energy content of prey, handling times of the predator on prey, space clearance rate i.e. how fast a predator searches space, and prey densities. Each of these variables is derived from the allometric scaling relationship using the body sizes of species. More details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in @petchey2008.

@@

The observed food webs that we fit the ADBM to belong to marine, freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). These food webs contain primary producers, herbivores, carnivores, parasites, and parasitoids and also contain various types of feeding interactions, including predation, herbivory, bacterivory, parasitism and pathogenic. The observed connectance of these food webs varies from 0.03 to 0.24 and the number of species varies from 29 to 239 species.

**OP: reduce the width of the final column of table 1.**
**OP: state how and why these set of 12 food webs were chosen to study by us.**

The goodness of fit of the ADBM's predictions depends on the interaction types in the food webs. Those with size-structured interactions, such as herbivory in aquatic ecosystems interactions being better predicted when compared to less size-structured ones such as parasitoids and terrestrial herbivory ones [@petchey2008].

```{r echo=FALSE, warning=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
fw_table <- read_excel("table/fw_table.xlsx")
kable(fw_table, format = "latex",
      caption = "\\label{fig:tab_1}Information about the food webs predicted using the ADBM.", escape = FALSE, align = "l") %>%
  kable_styling(font_size = 7, latex_options="scale_down", row_label_position = 1) %>%
  landscape(margin = "1cm") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "8em") %>%
  column_spec(7, width = "8em") %>%
  column_spec(8, width = "8em") %>%
  column_spec(9, width = "8em")

@@

The ADBM was parameterised using approximate Bayesian computation (ABC) where a set of parameter values were sampled from the prior distributions. That set of parameter values was either accepted or rejected based on how close the predicted food web is to the observed food web using an accuracy metric -- true skill statistic (TSS). The accepted parameter values then formed a posterior distribution. Further, prediction intervals of the true skill statistic and connectance of the predicted food webs were computed. In our study, we considered food webs where the predicted connectance lay within the 95% prediction interval. A detailed explanation of the parameterisation method can be found in @guptaSimultaneouslyEstimatingFood2022.

@@

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of the three criteria: removal of (i) the most-connected species, (ii) the least-connected species and (iii) randomly chosen species. The most-connected and least-connected criteria are based on the degree (i.e. the total number of links to resources and from consumers) of species. We considered these three criteria because the random extinction scenario takes into account all the theoretically possible extinction sequences of species that can occur in a food web and the extinction of most-connected species and least-connected species takes into account the two opposite extreme scenarios. These extinction scenarios have been widely used in studying species extinctions and collapse of food webs and other networks [@dunne2002network; @soleComplexityFragilityEcological2001; @dunne2004; @dunneCascadingExtinctionsCommunity2009; @albertStatisticalMechanicsComplex2002].

Given a primary removal of species in a food web, if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. Once no more secondary extinctions occurred, then another primary extinction was made, of the next appropriate species depending on the extinction scenario. This process was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of species subjected to primary removals that resulted in extinction (primary and secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, which results in at least 50% of total species loss [@dunne2002network; @dunne2004; @jonssonReliabilityR50Measure2015; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$), the first primary extinction causes a cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

@@

First, we simulated primary species loss in food webs predicted by the ADBM which had the maximum true skill statistics and compared it to primary species loss in observed food webs. Second, to take into account the uncertainty in robustness in the ADBM predicted food webs we simulated primary species loss and thereby computed robustness for all the ADBM predicted food webs corresponding to the 95% prediction interval of the predicted connectance. Furthermore, in the case of the random extinction scenario, we simulated 1000 random extinction sequences in a single ADBM predicted food web.

@@

In the random extinction scenario, we computed robustness $R_{50}$ for all 1000 independent random extinction sequences and calculated the median as a summary statistics to quantify the average robustness of a single food web to random extinction. To quantify the effect of undersampling, i.e. overestimation of connectance, we compute the ratio of the difference in normalised robustness between the ADBM predicted food webs and observed food webs to the difference in their normalised connectance, where normalisation was performed by dividing the variables by their maximum possible values (I.e 0.5 for $R_{50}$ and 1 for connectance). However, we did not perform any statistical significance test because we work with simulated food webs and therefore the p-values of these tests are influenced by the number of model simulations [@whiteEcologistsShouldNot2014].

@@

We first present the secondary extinction curves of the ADBM predicted food webs which had the maximum true skill statistics and of the observed food webs for 12 food webs under three different extinction scenarios. We then compare the robustness of all the ADBM predicted food webs within the 95% prediction interval against that of the observed food webs to take into account uncertainty in the robustness across food webs predictions. Finally we quantify the effect of overestimation of connectance on the difference in their robustness estimates.

@@

In the most-connected extinction scenario, the cumulative secondary extinction curves started to rise steeply at a lower number of primary species removal in the observed food webs as compared to the ADBM predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d, e, f, g, h, j, l)). However there were higher number of cumulative secondary extinctions occurring in the ADBM predicted food webs when compared to that of the observed food webs at a high number of primary species removal in five food webs (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith Pond food web, there were no secondary extinctions for any number of primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the Broadstone Stream (taxonomic aggregation) food web the same was true only for the observed food web but in the ADBM predicted food web there was a steep rise in the cumulative secondary extinctions (Fig. \ref{fig:fig_r1} (b)).

@@

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary extinctions in the ADBM predicted and the observed food webs for five (out of 1000) independent random extinction sequences to show example variation caused by different random primary extinction orders. The secondary extinctions in the ADBM predicted food webs were more abrupt than that in the observed food webs.

**OP: expand and clarify what is meant by abrupt... describe this in a sentence. (Abrupt could mean different things to different people, so clarification would be good.)

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary removals of \\textbf{random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_ra_extlines_maxTSS.png")
```

Compared to the most-connected and random extinction scenarios, there were fewer secondary extinctions in the least-connected extinction scenario and therefore the secondary extinction curves were flat for many of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs, the extinction curves of the ADBM predicted food webs overlapped with the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)}). In five of the food webs, a very high number of secondary extinctions occurred at very low number of primary species removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

@@

The ADBM predicted food webs were more robust than the observed food webs on average in the most-connected and random extinction scenarios (Fig. \ref{fig:fig_r4} (a, b)). However, there were large variations in the robustness within the ADBM predicted food webs in the most-connected extinction scenario (Fig. \ref{fig:fig_r4} (a)). For example, the ADBM predicted Caricaie Lakes food web was more robust than the observed food web on average but had a larger variation in the robustness within the ADBM predicted food webs compared to other food webs.

The food webs were more robust to the random extinction scenario than the most-connected scenario (Fig. \ref{fig:fig_r4} (a, b)). Small Reef and Benguela Pelagic food webs had more variations in robustness within the ADBM predicted food webs as compared to the other food webs (Fig. \ref{fig:fig_r4} (b)). Skipwith Pond, Broadstone Stream (taxonomic aggregation) and Broadstone Stream (size aggregation) food webs were the most robust (Median $R_{50} = 0.5$) for both ADBM predicted and observed food webs. Although there were few less robust ADBM predicted food webs in the Broadstone Stream (size aggregation) as shown by the outliers.

In the least-connected extinction scenario, the food webs had a very high robustness (Median $R_{50} = 0.5$) for most of the food webs (Fig. \ref{fig:fig_r4} (c)), however there were some exceptions. The ADBM predicted food webs for Small Reef and Benguela Pelagic had very low median robustness. Benguela Pelagic, Broom and Capinteria food webs from the ADBM had larger variations in robustness when compared to that of the others.

```{r, echo=FALSE, fig.align='center', out.width="450px", fig.cap="\\label{fig:fig_r4} Robustness comparison between the ADBM predicted food webs and the observed food webs for 12 food webs across different ecosystems. Here, $R_{50}$ is the proportion of species that have to be removed to achieve a total loss of at least 50\\% of total species (primary removals and secondary extinctions). Box represent 25th and 75th percentile; solid diamond represent median; whisker represent outlier limits; the outlier coefficient used was 1.5. Some points are not visible due to perfect overlap in b and c. Refer to Fig. 7 in the Supplementary Information for a faceted visualisation."}
knitr::include_graphics("../results/plot_R50_ADBM_vs_obs.png")
```

**OP: "Box" in the figure caption is stated, but it is not clear where boxes are in the figure.

>>>>

 1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occurring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that model predicted food webs usually have greater connectance than observed ones. Thus, we expect that predicted food webs are more robust than observed ones. However, this expectation has never, to our knowledge, been tested, nor has the effect size, i.e. the difference in robustness of predicted and observed food web with respect to the difference in their connectance, been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links but does not predict all links correctly. We did this for 12 different food webs from a wide variety of ecosystems. We used three extinction scenarios: random, most connected, and least connected.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, which can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found that the food webs robustness increased by 0.52 units for the most connected species extinction scenario and by 0.04 units for the random species extinction scenarios. On the other hand, we saw no effect, on average, of increased connectance for the least connected species extinction scenario for all except two of the 12 food webs.
  5) These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, so it would be interesting to compare the dynamical stability of observed and predicted food webs and the topological stability that we have focused on.

keywords:
  - connectance
  - ABC
  - ADBM
  - food web
  - extinction
  - uncertainty
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage {amsmath}
  - \setlength\parindent{24pt}
  - \usepackage{setspace}\doublespacing
# indent: true
mainfont: Times New Roman
bibliography: bibliography.bib
csl: methods-in-ecology-and-evolution.csl
biblio-style: biblatex
output:
  rticles::arxiv_article:
    keep_tex: true
---

@@

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity and can lead to food web collapse [@ullah2018]. This food web collapse is due to the cascades of secondary extinctions in a food web because of the primary loss of species due to environmental change [@pimmHumanImpactsRates2006; @thomasComparativeLossesBritish2004; @thomasExtinctionRiskClimate2004]. An example of a secondary extinction is when a consumer goes extinct because its sole resource species goes extinct. Therefore, research focused on cascading secondary extinctions, also known as 'community viability analysis', has been performed extensively in the past few decades to quantify how robust food webs are to species extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005]. This research revealed that the rate of collapse of a food web depends on its structure and complexity [@dunne2002network; @dunneCascadingExtinctionsCommunity2009].

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems, where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. Primary species loss is considered to be extinction of a species due to causes external to the food web, e.g. overharvesting [@koningGrassrootsReservesRescue2021], introduction of invasive species [@davidImpactsInvasiveSpecies2017]. A secondary extinction is an extinction caused by a previous extinction (either primary or secondary). Numerous studies have used topological criterion for assigning a secondary extinction to a species, i.e., if extinction is of the last resource species of a consumer species, then the consumer species suffers secondary extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @soleComplexityFragilityEcological2001].

When there are few primary extinctions, and these cause many secondary extinctions, a food web is said to be fragile or not robust. In contrast, a food web is said to be robust when there are only a few secondary extinctions. Several studies have shown that the robustness of the food webs increases with food web connectance [@dunneNetworkStructureBiodiversity2002; @dunneCascadingExtinctionsCommunity2009]. Additionally, these studies have demonstrated that the removal (primary extinction) of the most connected species causes considerably more secondary extinctions than the random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. Simulation studies like these, which investigate the impact of primary extinctions in a food web to quantify robustness based on its topological structure, provide an alternate solution to canonical experiments in natural ecosystems, which are not possible or very difficult to conduct [@dunneCascadingExtinctionsCommunity2009].

Along with quantifying food web robustness based on its topological structure, studies such as @williamsEffectsNetworkDynamical2008, @broseAllometricScalingEnhances2006, and @martinezDiversityComplexityPersistence have quantified robustness based on the abundance dynamics of a food web. However, the topological approach of quantifying a food web's robustness only requires the food web structure, whereas the dynamical approach requires the food web structure and the temporal dynamics of the abundance of species in that food web. For example, @williamsEffectsNetworkDynamical2008 combined network structure models with bioenergetic dynamics models to study the role of food web topology and nonlinear dynamics on species coexistence in complex ecological networks.

@@

A key assumption of the observed food webs is that they are very well sampled, i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and do not represent all of the feeding links that can occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort than others, whereas some trophic links remain unobserved because of biological constraints such as spatio-temporal uncoupling, size or reward mismatching, foraging constraints and physiological-biochemical constraints, which are irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. Previous studies such as @caronAddressingEltonianShortfall and @guptaSimultaneouslyEstimatingFood2022 have shown that the predicted food webs from these models usually have greater connectance than the observed ones. One solution to compensate for undersampling is using a food web model such as the Allometric Diet Breadth Model (ADBM) [@petchey2008] to predict the missing links and then measure the robustness of the predicted food web. The ADBM is a mechanistic model constructed using foraging rules based on the body sizes of prey and predator, where trophic interactions satisfying those rules would be predicted by the model, which are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links and also may predict some links that could not, in fact occur.

@@

In our study, we investigate the topological robustness of the ADBM predicted food webs and compare it to that of the observed food webs. We expect the ADBM predicted food webs to be more robust than the observed food webs, and for the greater robustness to be related to the amount by which the ADBM predicts greater connectance compared to that of the observed food webs. We do this by simulating primary species loss in 12 food webs predicted by the ADBM to quantify the secondary loss of extinctions and compare the robustness of food webs predicted by the ADBM to the robustness of the observed food webs. We use three different approaches of primary species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the implementation of the simulation of primary extinctions for three different extinction scenarios on 12 food webs predicted by the ADBM from a wide variety of ecosystems and compute the resultant secondary extinctions. We then describe a robustness metric of those predicted food webs and how we compare them to the properties of the food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. We chose this model because it is known to predict greater connectance in the predicted food webs compared to the observed food webs [@guptaSimultaneouslyEstimatingFood2022]. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables used in the model are the energy content of prey, handling times of the predator on prey, space clearance rate, i.e. how fast a predator searches space, and prey densities. Each of these variables is derived from the allometric scaling relationship using the body sizes of species. More details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in open access @petchey2008.

@@

The observed food webs that we fit the ADBM to belong to marine, freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). We considered these food webs because they belong to diverse ecosystems and follow FAIR (Findable Accessible Interoperable Reusable) principles [@wilkinsonFAIRGuidingPrinciples2016]. These food webs contain primary producers, herbivores, carnivores, parasites, and parasitoids and various feeding interactions, including predation, herbivory, bacterivory, parasitism and pathogenic. The observed connectance of these food webs varies from 0.03 to 0.24 (Links/Species$^2$), and the number of species varies from 29 to 239. The goodness of fit of the ADBM's predictions depends on the interaction types in the food webs. For example, those with size-structured interactions, such as herbivory in aquatic ecosystems, are better predicted compared to less size-structured ones, such as parasitoids and terrestrial herbivory ones [@petchey2008].

```{r echo=FALSE, warning=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
fw_table <- read_excel("table/fw_table.xlsx")
kable(fw_table, format = "latex",
      caption = "\\label{fig:tab_1}Information about the food webs predicted using the ADBM.", escape = FALSE, align = "l") %>%
  kable_styling(font_size = 7, latex_options="scale_down", row_label_position = 1) %>%
  landscape(margin = "1cm") %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "3cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "2cm") %>%
  column_spec(6, width = "3cm")

@@

The ADBM was parameterised using approximate Bayesian computation (ABC), where a set of parameter values were sampled from a prior distribution. Then, that set of parameter values was either accepted or rejected based on how close the predicted food web was to the observed food web using an accuracy metric -- true skill statistic (TSS). The accepted parameter values formed a posterior distribution (Fig. 4 and S14-S25 in @guptaSimultaneouslyEstimatingFood2022). Further, prediction intervals of the true skill statistic and connectance of the predicted food webs were computed (Fig. 5 (a, b) in @guptaSimultaneouslyEstimatingFood2022). In our study, we considered model predicted food webs where the predicted connectance lay within the 95% prediction interval of all model predicted food webs. A detailed explanation of the parameterisation method can be found in @guptaSimultaneouslyEstimatingFood2022.

@@

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of the three criteria: removal of (i) the most connected species, (ii) the least connected species and (iii) randomly chosen species. The most connected and least connected criteria are based on species' degree (i.e. the total number of links to resources and from consumers). We considered these three criteria because the random extinction scenario takes into account all the theoretically possible extinction sequences of species that can occur in a food web, while the extinction of the most connected species and least connected species takes into account the two opposite extreme scenarios. These extinction scenarios have been widely used in studying species extinctions and the collapse of food webs and other networks [@dunne2002network; @soleComplexityFragilityEcological2001; @dunne2004; @dunneCascadingExtinctionsCommunity2009; @albertStatisticalMechanicsComplex2002].

Given a primary removal of species in a food web, if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web, and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. Once no more secondary extinctions occurred, then another primary extinction was made of the next appropriate species depending on the extinction scenario. This process was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of species subjected to primary removals resulting in extinction (primary and secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, which results in at least 50% of total species loss [@dunne2002network; @dunne2004; @jonssonReliabilityR50Measure2015; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$, where $S$ is the number of species), the first primary extinction causes a cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

@@

First, we simulated primary species loss in food webs predicted by the ADBM which had the maximum true skill statistics and compared it to primary species loss in observed food webs. Second, to take into account the uncertainty in robustness in the ADBM predicted food webs, we simulated primary species loss and thereby computed robustness for all the ADBM predicted food webs corresponding to the 95% prediction interval of the predicted connectance. Furthermore, in the case of the random extinction scenario, we simulated 1000 random extinction sequences in a single ADBM predicted food web.

@@

In the random extinction scenario, we computed robustness $R_{50}$ for all 1000 independent random extinction sequences and calculated the median as a summary statistics to quantify the average robustness of a single food web to random extinction. To quantify the effect of undersampling, i.e. greater connectance of connectance, we compute the ratio of the difference in normalised robustness between the ADBM predicted food webs and observed food webs to the difference in their normalised connectance, where normalisation was performed by dividing the variables by their maximum possible values (i.e. 0.5 for $R_{50}$ and 1 for connectance). However, we did not perform any statistical significance test because we work with simulated food webs and therefore, the p-values of these tests are influenced by the number of model simulations [@whiteEcologistsShouldNot2014].

@@

We first present the secondary extinction curves of the ADBM predicted food webs which had the maximum true skill statistics and of the observed food webs for 12 food webs under three different extinction scenarios. We then compare the robustness of all the ADBM predicted food webs within the 95% prediction interval against that of the observed food webs to take into account uncertainty in the robustness across food webs predictions. Finally, we quantify the effect of the difference in their connectance on the difference in their robustness estimates.

@@

In the most connected extinction scenario, the cumulative secondary extinction curves started to rise steeply in the observed food webs at a lower number of primary species removal as compared to the ADBM predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d, e, f, g, h, j, l)). However, there was a higher number of cumulative secondary extinctions occurring in the ADBM predicted food webs when compared to that of the observed food webs at a high number of primary species removal in five food webs (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith Pond food web, there were no secondary extinctions for any number of primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the Broadstone Stream (taxonomic aggregation) food web the same was true only for the observed food web but in the ADBM predicted food web there was a steep rise in the cumulative secondary extinctions (Fig. \ref{fig:fig_r1} (b)).

@@

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary extinctions in the ADBM predicted and the observed food webs for five (out of 1000) independent random extinction sequences to show example variation caused by different random primary extinction orders. The secondary extinction curves of the ADBM predicted food webs were steeper than that of the observed food webs, i.e. primary removal of some species in an extinction sequence can lead to the complete collapse of the remaining food web in the ADBM predicted food webs.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary removals of \\textbf{random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five (out ot 1000) independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_ra_extlines_maxTSS.png")
```

Compared to the most connected and random extinction scenarios, there were fewer secondary extinctions in the least connected extinction scenario, and therefore the secondary extinction curves were flat for many of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs, the extinction curves of the ADBM predicted food webs overlapped with the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)}). In five of the food webs, a very high number of secondary extinctions occurred at a very low number of primary species removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

@@

The ADBM predicted food webs were more robust than the observed food webs on average in the most connected and random extinction scenarios (Fig. \ref{fig:fig_r4} (a, b)). However, there were large variations in the robustness within the ADBM predicted food webs in the most connected extinction scenario (Fig. \ref{fig:fig_r4} (a)). For example, the ADBM predicted Caricaie Lakes food web was more robust than the observed food web on average but had a larger variation in the robustness within the ADBM predicted food webs compared to other food webs.

The food webs were more robust to the random extinction scenario than the most connected scenario (Fig. \ref{fig:fig_r4} (a, b)). Small Reef and Benguela Pelagic food webs had more variations in robustness within the ADBM predicted food webs as compared to the other food webs (Fig. \ref{fig:fig_r4} (b)). Skipwith Pond, Broadstone Stream (taxonomic aggregation) and Broadstone Stream (size aggregation) food webs were the most robust (Median $R_{50} = 0.5$) for both ADBM predicted and observed food webs. Although there were few less robust ADBM predicted food webs in the Broadstone Stream (size aggregation) as shown by the outliers. The food webs were more robust to the random extinction scenario than the most connected scenario (Fig. \ref{fig:fig_r4} (a, b)). Small Reef and Benguela Pelagic food webs had more variations in robustness within the ADBM predicted food webs as compared to the other food webs (Fig. \ref{fig:fig_r4} (b)). Skipwith Pond, Broadstone Stream (taxonomic aggregation) and Broadstone Stream (size aggregation) food webs were the most robust (Median $R_{50} = 0.5$) for both ADBM predicted and observed food webs. Although there were few less robust ADBM predicted food webs in the Broadstone Stream (size aggregation) as shown by the outliers.

In the least connected extinction scenario, the food webs had a very high robustness (Median $R_{50} = 0.5$) for most of the food webs (Fig. \ref{fig:fig_r4} (c)), however there were some exceptions. The ADBM predicted food webs for Small Reef and Benguela Pelagic had very low median robustness. Benguela Pelagic, Broom and Capinteria food webs from the ADBM had larger variations in robustness when compared to that of the others.

```{r, echo=FALSE, fig.align='center', out.width="450px", fig.cap="\\label{fig:fig_r4} Robustness comparison between the ADBM predicted food webs and the observed food webs for 12 food webs across different ecosystems. Here, $R_{50}$ is the proportion of species that have to be removed to achieve a total loss of at least 50\\% of total species (primary removals and secondary extinctions). Box represent the 25th and 75th percentile; solid diamond represents the median; whisker represents outlier limits; the outlier coefficient used was 1.5. Some points are not visible due to perfect overlap in b and c. Refer to Fig. 7 in the Supplementary Information for a faceted visualisation. The dashed black lines are the 1:1 relationships for reference."}
knitr::include_graphics("../results/plot_R50_ADBM_vs_obs.png")
```
